An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-02-12 EconGrowth.Rnw
Added better variables for graph widths and heights in the setup chunk.

<<setup, echo=FALSE, message=FALSE>>=
require(knitr)
require(lattice)
require(latticeExtra)
require(ggplot2)
require(car)
require(mosaic)
require(xtable)
# tikz allows use of LaTeX formatting and font in graphs. Allows for a consistent look across the paper.
# See http://r-forge.r-project.org/R/?group_id=440 for instructions on installing tikzDevice.
require(tikzDevice) 
trellis.par.set(theme=theme.mosaic())
options(width=75)
options(show.signif.stars=FALSE)
opts_chunk$set(
  # ********************************************************************************************************
  eval=FALSE,
  # ********************************************************************************************************
  warning=FALSE, #Tells whether to show warnings in the output.
  dev='tikz',    #Allows LaTeX code in graphical output. E.g., "$y$" for a variable name in the legend for a graph.
  tidy=FALSE,
  comment=NA,
  #Tells whether to cache output from chunks are saved. Cacheing saves time. 
  #However, references to LaTeX tables DO NOT WORK if cacheing is turned on (cache=TRUE).
  cache=FALSE,
  echo=FALSE,    #Tells whether to echo code for all chunks. Set TRUE on an individual chunk to see its code.
  message=FALSE) #Turns off messages for all chunks. Set TRUE on an individual chunk to see it.

# Statistical significance levels. We'll work with 95% CIs
ciLevel <- 0.95
ciHalfLevel <- ciLevel + (1.0-ciLevel)/2.0
# List of countries
countryAbbrevs <- c(US="US", UK="UK", JP="JP", CN="CN", ZA="ZA", SA="SA", IR="IR", TZ="TZ", ZM="ZM")
countryAbbrevsU <- c(US="US", UK="UK", JP="JP") #Only these countries have useful work data
countryNamesAlph <- c(CN="China", IR="Iran", JP="Japan", SA="Saudi Arabia", TZ="Tanzania", UK="United Kingdom", US="USA", ZA="South Africa", ZM="Zambia") #In alphabetical order.
energyTypes <- c(Q="Q", X="X", U="U") # List of energy types
factors <- c(K="K", L="L", Q="Q", X="X", U="U") # List of factors of production
########### Several global parameters for graphs. Set here and use below to ensure consistent appearance of graphs.
# Set the order for presenting countries in 3x3 lattice graphs. Default is alphabetical. 
# "1" means first alphabetically.
countryOrderForGraphs <- c(7,6,3,1,8,5,4,2,9) # Sets the order as US, UK, JP, CN, ZA, TZ, SA, IR, ZM.
# Full page lattice plot sizes and other parameters
ninePanelLayoutSpec <- c(3,3) #indicates a 3x3 arrangement of panels.
figWidth <- 6.5             #Inches
ninePanelGraphHeight <- 7.5 #Inches
parameterGraphHeight <- 5   #Inches
keyTextSize <- 0.85 #85% of normal size
keyColumns <- 1 #Want only 1 column in the key for lattice graphs
keyXLoc <- 0.01 #x position of the key
keyYLoc <- 0.95 #y position of the key
scaleTextSize <- 1.0  #Multiple of normal size
scaleTickSize <- -0.5 #50% of normal size and pointing INWARD!
@

@@

<<Factors_Lattice_Graph, fig.width=figWidth, fig.height=ninePanelGraphHeight, fig.pos="H", fig.cap="Historical data. Indexed GDP ($y$), capital stock ($k$), labor ($l$), thermal energy ($q$), exergy ($x$), and useful work ($u$).">>=
createFactorsLatticeGraph <- function(){
  dataTable <- loadData("All")
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1, 5, 2, 4, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 2, 2, 1, 1, 1) #line widths. 0 means no line.
  colors <- c("black", "black", "black", "red", "blue", "darkorange") #line and symbol colors
  symbols <- c(1, NA, NA, NA, NA, NA)  #NA gives no symbol.
  graph <- xyplot(iGDP+iCapStk+iLabor+iQ+iX+iU ~ Year | Country, data=dataTable,
                  type = graphType,
                  index.cond = list(countryOrderForGraphs), #orders the panels.
                  layout = ninePanelLayoutSpec, 
                  strip = strip.custom(factor.levels=countryNamesAlph, bg="white"), #Sets strip parameters
                  as.table = TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("$y$", "$k$", "$l$", "$q$", "$x$", "$u$")),
                           type=graphType,
                           cex=keyTextSize, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              y=list(relation="free", #allows each axis to be different
                                     at=c(1,2,3,4,5,6,7,8,9,10) #tick mark locations
                                     )
                              ),
                  ylim=list(c(1,10), c(1,4), c(1,4), c(1,4), c(1,4), c(1,4), c(1,4), c(1,4), c(1,4)), #y axis limits
                  xlab="", ylab="Indexed (1980=1 or 1991=1)") #axis labels
  return(graph)
}
createFactorsLatticeGraph()
@

@@

<<SFk_Params_Graph, fig.width=figWidth, fig.height=parameterGraphHeight, fig.pos="H", fig.cap="Single-factor (with $k$) model parameters. Vertical bars indicate 95\\% confidence intervals.", eval=TRUE>>=
createSFParamsGraph("K")
@

<<SFl_Params_Graph, fig.width=figWidth, fig.height=parameterGraphHeight, fig.pos="H", fig.cap="Single-factor (with $l$) model parameters. Vertical bars indicate 95\\% confidence intervals.", eval=TRUE>>=
createSFParamsGraph("L")
@

<<SFq_Params_Graph, fig.width=figWidth, fig.height=parameterGraphHeight, fig.pos="H", fig.cap="Single-factor (with $q$) model parameters. Vertical bars indicate 95\\% confidence intervals.", eval=TRUE>>=
createSFParamsGraph("Q")
@

<<SFx_Params_Graph, fig.width=figWidth, fig.height=parameterGraphHeight, fig.pos="H", fig.cap="Single-factor (with $x$) model parameters. Vertical bars indicate 95\\% confidence intervals.", eval=TRUE>>=
createSFParamsGraph("X")
@

<<SFu_Params_Graph, fig.width=figWidth, fig.height=parameterGraphHeight, fig.pos="H", fig.cap="Single-factor (with $u$) model parameters. Vertical bars indicate 95\\% confidence intervals.", eval=TRUE>>=
createSFParamsGraph("U")
@

@@

<<SF_GDP_Lattice_Graph, fig.width=figWidth, fig.height=ninePanelGraphHeight, fig.pos="H", fig.cap="Single-factor results.">>=
createSFLatticeGraph()
@

@@

<<CD_Params_Graph, fig.width=figWidth, fig.height=parameterGraphHeight, fig.pos="H", fig.cap="Cobb-Douglas (without energy) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCDParamsGraph()
@

@@

<<CDq_Params_Graph, fig.width=figWidth, fig.height=parameterGraphHeight, fig.pos="H", fig.cap="Cobb-Douglas (with $q$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCDParamsGraph("Q")
@

@@

<<CDx_Params_Graph, fig.width=figWidth, fig.height=parameterGraphHeight, fig.pos="H", fig.cap="Cobb-Douglas (with $x$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCDParamsGraph("X")
@

@@

<<CDu_Params_Graph, fig.width=figWidth, fig.height=parameterGraphHeight, fig.pos="H", fig.cap="Cobb-Douglas (with $u$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCDParamsGraph("U")
@

@@

<<CD_GDP_Lattice_Graph, fig.width=figWidth, fig.height=ninePanelGraphHeight, fig.pos="H", fig.cap="Cobb-Douglas results.">>=
createCDLatticeGraph()
@

>>>>

<<setup, echo=FALSE, message=FALSE>>=
require(knitr)
require(lattice)
require(latticeExtra)
require(ggplot2)
require(car)
require(mosaic)
require(xtable)
# tikz allows use of LaTeX formatting and font in graphs. Allows for a consistent look across the paper.
# See http://r-forge.r-project.org/R/?group_id=440 for instructions on installing tikzDevice.
require(tikzDevice) 
trellis.par.set(theme=theme.mosaic())
options(width=75)
options(show.signif.stars=FALSE)
opts_chunk$set(
  # ********************************************************************************************************
  eval=TRUE,
  # ********************************************************************************************************
  warning=FALSE, #Tells whether to show warnings in the output.
  dev='tikz',    #Allows LaTeX code in graphical output. E.g., "$y$" for a variable name in the legend for a graph.
  tidy=FALSE,
  comment=NA,
  #Tells whether to cache output from chunks are saved. Cacheing saves time. 
  #However, references to LaTeX tables DO NOT WORK if cacheing is turned on (cache=TRUE).
  cache=FALSE,
  echo=FALSE,    #Tells whether to echo code for all chunks. Set TRUE on an individual chunk to see its code.
  message=FALSE) #Turns off messages for all chunks. Set TRUE on an individual chunk to see it.

# Statistical significance levels. We'll work with 95% CIs
ciLevel <- 0.95
ciHalfLevel <- ciLevel + (1.0-ciLevel)/2.0
# List of countries
countryAbbrevs <- c(US="US", UK="UK", JP="JP", CN="CN", ZA="ZA", SA="SA", IR="IR", TZ="TZ", ZM="ZM")
countryAbbrevsU <- c(US="US", UK="UK", JP="JP") #Only these countries have useful work data
countryNamesAlph <- c(CN="China", IR="Iran", JP="Japan", SA="Saudi Arabia", TZ="Tanzania", UK="United Kingdom", US="USA", ZA="South Africa", ZM="Zambia") #In alphabetical order.
energyTypes <- c(Q="Q", X="X", U="U") # List of energy types
factors <- c(K="K", L="L", Q="Q", X="X", U="U") # List of factors of production
########### Several global parameters for graphs. Set here and use below to ensure consistent appearance of graphs.
# Set the order for presenting countries in 3x3 lattice graphs. Default is alphabetical. 
# "1" means first alphabetically.
countryOrderForGraphs <- c(7,6,3,1,8,5,4,2,9) # Sets the order as US, UK, JP, CN, ZA, TZ, SA, IR, ZM.
# Full page lattice plot sizes and other parameters
ninePanelLayoutSpec <- c(3,3) #indicates a 3x3 arrangement of panels.
ninePanelGraphWidth <- 6.5   #Inches
ninePanelGraphHeight <- 7.5  #Inches
sfParameterGraphWidth <- 4   #Inches
sfParameterGraphHeight <- 3  #Inches
cdParameterGraphWidth <- 6.5 #Inches
cdParameterGraphHeight <- 5  #Inches
keyTextSize <- 0.85 #85% of normal size
keyColumns <- 1 #Want only 1 column in the key for lattice graphs
keyXLoc <- 0.01 #x position of the key
keyYLoc <- 0.95 #y position of the key
scaleTextSize <- 1.0  #Multiple of normal size
scaleTickSize <- -0.5 #50% of normal size and pointing INWARD!
@

@@

<<Factors_Lattice_Graph, fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.pos="H", fig.cap="Historical data. Indexed GDP ($y$), capital stock ($k$), labor ($l$), thermal energy ($q$), exergy ($x$), and useful work ($u$).">>=
createFactorsLatticeGraph <- function(){
  dataTable <- loadData("All")
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1, 5, 2, 4, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 2, 2, 1, 1, 1) #line widths. 0 means no line.
  colors <- c("black", "black", "black", "red", "blue", "darkorange") #line and symbol colors
  symbols <- c(1, NA, NA, NA, NA, NA)  #NA gives no symbol.
  graph <- xyplot(iGDP+iCapStk+iLabor+iQ+iX+iU ~ Year | Country, data=dataTable,
                  type = graphType,
                  index.cond = list(countryOrderForGraphs), #orders the panels.
                  layout = ninePanelLayoutSpec, 
                  strip = strip.custom(factor.levels=countryNamesAlph, bg="white"), #Sets strip parameters
                  as.table = TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("$y$", "$k$", "$l$", "$q$", "$x$", "$u$")),
                           type=graphType,
                           cex=keyTextSize, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              y=list(relation="free", #allows each axis to be different
                                     at=c(1,2,3,4,5,6,7,8,9,10) #tick mark locations
                                     )
                              ),
                  ylim=list(c(1,10), c(1,4), c(1,4), c(1,4), c(1,4), c(1,4), c(1,4), c(1,4), c(1,4)), #y axis limits
                  xlab="", ylab="Indexed (1980=1 or 1991=1)") #axis labels
  return(graph)
}
createFactorsLatticeGraph()
@

@@

<<SFk_Params_Graph, fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.pos="H", fig.cap="Single-factor (with $k$) model parameters. Vertical bars indicate 95\\% confidence intervals.", eval=TRUE>>=
createSFParamsGraph("K")
@

<<SFl_Params_Graph, fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.pos="H", fig.cap="Single-factor (with $l$) model parameters. Vertical bars indicate 95\\% confidence intervals.", eval=TRUE>>=
createSFParamsGraph("L")
@

<<SFq_Params_Graph, fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.pos="H", fig.cap="Single-factor (with $q$) model parameters. Vertical bars indicate 95\\% confidence intervals.", eval=TRUE>>=
createSFParamsGraph("Q")
@

<<SFx_Params_Graph, fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.pos="H", fig.cap="Single-factor (with $x$) model parameters. Vertical bars indicate 95\\% confidence intervals.", eval=TRUE>>=
createSFParamsGraph("X")
@

<<SFu_Params_Graph, fig.width=sfParameterGraphWidth, fig.height=sfParameterGraphHeight, fig.pos="H", fig.cap="Single-factor (with $u$) model parameters. Vertical bars indicate 95\\% confidence intervals.", eval=TRUE>>=
createSFParamsGraph("U")
@

@@

<<SF_GDP_Lattice_Graph, fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.pos="H", fig.cap="Single-factor results.">>=
createSFLatticeGraph()
@

@@

<<CD_Params_Graph, fig.width=cdParameterGraphWidth, fig.height=cdParameterGraphHeight, fig.pos="H", fig.cap="Cobb-Douglas (without energy) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCDParamsGraph()
@

@@

<<CDq_Params_Graph, fig.width=cdParameterGraphWidth, fig.height=cdParameterGraphHeight, fig.pos="H", fig.cap="Cobb-Douglas (with $q$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCDParamsGraph("Q")
@

@@

<<CDx_Params_Graph, fig.width=cdParameterGraphWidth, fig.height=cdParameterGraphHeight, fig.pos="H", fig.cap="Cobb-Douglas (with $x$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCDParamsGraph("X")
@

@@

<<CDu_Params_Graph, fig.width=cdParameterGraphWidth, fig.height=cdParameterGraphHeight, fig.pos="H", fig.cap="Cobb-Douglas (with $u$) model parameters. Vertical bars indicate 95\\% confidence intervals.">>=
createCDParamsGraph("U")
@

@@

<<CD_GDP_Lattice_Graph, fig.width=cdParameterGraphWidth, fig.height=ninePanelGraphHeight, fig.pos="H", fig.cap="Cobb-Douglas results.">>=
createCDLatticeGraph()
@
